---
title: "avengersLPS"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Introduction

This is a report on an analysis performed on a dataset that details the deaths (and subsequent return to life) of Marvel comic book superheroes that have joined the Avengers team, between the time they have joined and April of 2015. This dataset has been provided by an Internet website called FiveThirtyEight and is organized as a list of all members of the Avengers with some of their characteristics. For each Avenger there are binary variables that inform if the hero has died one, two, three, four or five times in the past. The same information is available if the hero has returned to life, from one to five times as well. 

We analyze this dataset through three research questions:

* Is there some characteristic that makes an Avenger more likely to die?
* Is there some characteristic that makes an Avenger more likely to revive?
* Is the number of comic book appearances of a superhero correlated to the number of years since he joined the team?


#Loading the CSV File and preprocessing the dataframe

Firstly, we download the CSV file from the FiveThirtyEight Github and save it as "avengers.csv"

```{r}
library(tidyverse)
file = "avengers.csv"
if(!file.exists(file)){
  download.file("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv",
	destfile=file)
}
```

Then, we load the file "avengers.csv" in the dataframe "avengers_df"
```{r}
library(readr);
avengers_df <- read.csv("avengers.csv", quote="\"");
head(avengers_df)
```

##Changing the type of some columns

Based on an overall analysis of the data, we noticed that some columns have inadequate typing, specifically "URL", "Name.Alias" and "Notes" , which should be string. So, we convert them:

```{r}
avengers_df$URL <-  as.character(avengers_df$URL)
avengers_df$Name.Alias <-  as.character(avengers_df$Name.Alias)
avengers_df$Notes <-  as.character(avengers_df$Notes)
```



##Converting to Tidy data

As mentioned in the introduction, the dataset contains five columns that binarily register if the Avenger has died. Each column represents their first, second, third, fourth and fifth deaths. Similarly, there are also five columns that represent if the Avenger has returned to life. These 10 columns simply contain YES or NO values.

There is no information of when the Avenger has died. In fact, the dataset is essentially a list of all Avengers that have ever existed. Therefore, converting each death into different rows does not make sense.

Instead, to make the number of deaths and revivals simpler, we will convert the number of YES values for each variable into an integer.


```{r}
library(tidyverse)

avengers_tidydf <- avengers_df %>% 
  mutate(Total.Deaths = avengers_df %>% select(Death1,Death2,Death3,Death4,Death5) %>% apply(1, function(x) length(x[x=="YES"])))   %>% 
  mutate(Total.Return = avengers_df %>% select(Return1,Return2,Return3,Return4,Return5) %>% apply(1, function(x) length(x[x=="YES"]))) %>%
  subset(select=-c(Death1,Death2,Death3,Death4,Death5,Return1,Return2,Return3,Return4,Return5))
head(avengers_tidydf)


```

#Question 1: Is there some characteristic that makes an Avenger more likely to have died?

We explored this question based on four characteristics available in the dataset: Gender, honorary status, number of years since joining the Avengers, and the number of comic book appearances of the hero. The plots in this section share a common presentaion method, which is the normalized stacked plot. This makes it easier to compare two groups of Avengers through diferences in the ratio of the total number of deaths in that group.

##Gender

Let's start with gender. From the graph, we can see that women are more likely to not die than men (64% to 58%) compared to dying exactly one time (26% to 33%). 


```{r}
  avengers_tidydf %>% 
  select(Gender,Total.Deaths) %>% 
  group_by(Gender,Total.Deaths) %>%
  summarize(NumberOfAvengers = n()) %>%
  mutate(percent= paste0(round(100*NumberOfAvengers/sum(NumberOfAvengers)), '%')) %>%
  ggplot(aes(x= Gender, y = NumberOfAvengers, fill = as.factor(Total.Deaths))) +
  geom_col(position = "fill", width = .5) +
  scale_fill_brewer(name ="Total Number of Deaths", palette = "Dark2") +
  geom_text(aes(label= paste0(NumberOfAvengers, ' (',percent,')')), position = position_fill(vjust=.5)) +
  xlab("Gender") +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```
##Honorary Status

Let's try honorary status. As seen in the graph, Academic and Honorary Avengers are more likely to not have died compared to Full-time Avengers. Probationary has only two Avengers, so it is not possible to infer anything.

```{r}
  avengers_tidydf %>% 
  select(Honorary,Total.Deaths) %>% 
  group_by(Honorary,Total.Deaths) %>%
  summarize(NumberOfAvengers = n()) %>%
  mutate(percent= paste0(round(100*NumberOfAvengers/sum(NumberOfAvengers)), '%')) %>%
  ggplot(aes(x= Honorary, y = NumberOfAvengers, fill = as.factor(Total.Deaths))) +
  geom_col(position = "fill", width = .7) +
  scale_fill_brewer(name ="Total Number of Deaths", palette = "Dark2") +
  geom_text(aes(label= paste0(NumberOfAvengers, ' (',percent,')')), position = position_fill(vjust=.5)) +
  xlab("Honorary Status") +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```

##Number of Years since joining the Avengers


Let's analyze the Years.since.joining variable. Years.since.joining is equal to, by the datasets definition, the year 2015 (when the dataset was created) minus the year the character was introduced as a member of the Avengers. Because of this, there is some problematic values in these columns, in that the year the character was introduced is 1900. 

```{r}
  avengers_tidydf %>% 
  select(Year) %>% 
  group_by(Year) %>%  
  summarize(NumberOfAvengers = n()) %>%
  ggplot(aes(x= Year, y = NumberOfAvengers, )) +
  geom_col() +
  scale_x_continuous(name= "Year a character was introduced as a member of the Avengers") +
  scale_y_continuous(name = "Number of Avengers")#+
```


It is possible (and likely, looking at the Marvel wiki entries on these characters) that this value is simply the earliest possible year for the variable, which means that the year the character was introduced is unknown. Therefore, these plots should filter out these Avengers.

Since Years.since.joining is an integer variable, I grouped the rows by its 5 quantiles. Its subtle, but the longer an Avenger has been part of the team, the more likely it is for him to have died.

```{r}
  avengers_tidydf %>% 
  filter(Year != 1900) %>%
  select(Years.since.joining,Total.Deaths) %>%
  arrange(Years.since.joining) %>% 
  mutate(quartile = as.factor(ntile(Years.since.joining, 5))) %>%
  group_by(quartile,Total.Deaths) %>%
  summarize(NumberOfAvengers = n()) %>%
  ggplot(aes(x= quartile, y = NumberOfAvengers, fill = as.factor(Total.Deaths))) +
  geom_col(position = "fill", width = .5) +
  scale_fill_brewer(name ="Total Number of Deaths", palette = "Dark2") +
  geom_text(aes(label= NumberOfAvengers), position = position_fill(vjust=.5)) +
  scale_x_discrete(name= "Quantile" ,labels = c("First", "Second", "Third", "Fourth", "Fifth", "Sixth", "Seventh", "Eighth", "Ninth", "Tenth")) +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```

##Number of Appereances in comic books

I repeated the previous process for the number of appereances. The same thing can also be said. The more comic book appereances an Avenger has, the more likely it is for him to have died. This may imply that these variables have some correlation. We investigate this question in the last section of this report.


```{r}
  avengers_tidydf %>% 
  select(Appearances,Total.Deaths) %>% 
  arrange(Appearances) %>% 
  mutate(quartile = as.factor(ntile(Appearances, 5))) %>%
  group_by(quartile,Total.Deaths) %>%
  summarize(NumberOfAvengers = n()) %>%
  ggplot(aes(x= quartile, y = NumberOfAvengers, fill = as.factor(Total.Deaths))) +
  geom_col(position = "fill", width = .5) +
  scale_fill_brewer(name ="Total Number of Deaths", palette = "Dark2") +
  geom_text(aes(label= NumberOfAvengers), position = position_fill(vjust=.5)) +
  scale_x_discrete(name= "Quantile" ,labels = c("First", "Second", "Third", "Fourth", "Fifth", "Sixth", "Seventh", "Eighth", "Ninth", "Tenth")) +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```

#Question 2: Is there some characteristic that makes an Avenger more likely to have revived?

##Filtering all Avengers that have not died

The method for investigating why an Avenger may return to life is similar to the method of the previous section. Thus, most plots from question 1 are repeated in this section, though altered to investigate "Total.Returns" instead of "Total.Deaths". There is, however, one preprocessing step, in which we filter all Avengers that have not died from the dataset. We do so:

```{r}
avengersThatDied_tidydf <- avengers_tidydf %>%
                           filter(Total.Deaths!=0)
head(avengersThatDied_tidydf)
```


##Gender

Let's start with Gender. Similarly with deaths, women are more likely to have returned to life than men.

```{r}
  avengersThatDied_tidydf %>% 
  select(Gender,Total.Return) %>% 
  group_by(Gender,Total.Return) %>%
  summarize(NumberOfAvengers = n()) %>%
  mutate(percent= paste0(round(100*NumberOfAvengers/sum(NumberOfAvengers)), '%')) %>%
  ggplot(aes(x= Gender, y = NumberOfAvengers, fill = as.factor(Total.Return))) +
  geom_col(position = "fill", width = .5) +
  scale_fill_brewer(name ="Total Number of Revivals", palette = "Dark2") +
  geom_text(aes(label= paste0(NumberOfAvengers, ' (',percent,')')), position = position_fill(vjust=.5)) +
  xlab("Gender") +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```
##Honorary Status

In the case of honorary status, proportionaly FUll-Time Avengers are more likely 

```{r}
  avengersThatDied_tidydf %>% 
  select(Honorary,Total.Return) %>% 
  group_by(Honorary,Total.Return) %>%
  summarize(NumberOfAvengers = n()) %>%
  mutate(percent= paste0(round(100*NumberOfAvengers/sum(NumberOfAvengers)), '%')) %>%
  ggplot(aes(x= Honorary, y = NumberOfAvengers, fill = as.factor(Total.Return))) +
  geom_col(position = "fill", width = .5) +
  scale_fill_brewer(name ="Total Number of Revivals", palette = "Dark2") +
  geom_text(aes(label= paste0(NumberOfAvengers, ' (',percent,')')), position = position_fill(vjust=.5)) +
  xlab("Honorary Status") +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```

##Number of Years since joining the Avengers


```{r}
  avengersThatDied_tidydf %>%
  filter(Year != 1900) %>%
  select(Years.since.joining,Total.Return) %>% 
  arrange(Years.since.joining) %>% 
  mutate(quartile = as.factor(ntile(Years.since.joining, 10))) %>%
  group_by(quartile,Total.Return) %>%
  summarize(NumberOfAvengers = n()) %>%
  ggplot(aes(x= quartile, y = NumberOfAvengers, fill = as.factor(Total.Return))) +
  geom_col(position = "fill", width = .5) +
  scale_fill_brewer(name ="Total Number of Revivals", palette = "Dark2") +
  geom_text(aes(label= NumberOfAvengers), position = position_fill(vjust=.5)) +
  scale_x_discrete(name= "Quantile" ,labels = c("First", "Second", "Third", "Fourth", "Fifth", "Sixth", "Seventh", "Eighth", "Ninth", "Tenth")) +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```

##Number of Appereances in comic books

```{r}
  avengersThatDied_tidydf %>% 
  select(Appearances,Total.Return) %>% 
  arrange(Appearances) %>% 
  mutate(quartile = as.factor(ntile(Appearances, 10))) %>%
  group_by(quartile,Total.Return) %>%
  summarize(NumberOfAvengers = n()) %>%
  ggplot(aes(x= quartile, y = NumberOfAvengers, fill = as.factor(Total.Return))) +
  geom_col(position = "fill", width = .5) +
  scale_fill_brewer(name ="Total Number of Revivals", palette = "Dark2") +
  geom_text(aes(label= NumberOfAvengers), position = position_fill(vjust=.5)) +
  scale_x_discrete(name= "Quantile" ,labels = c("First", "Second", "Third", "Fourth", "Fifth", "Sixth", "Seventh", "Eighth", "Ninth", "Tenth")) +
  scale_y_continuous(name = "Number of Avengers", labels = scales::percent)#+
```

#Question 3: Is the number of comic book appearances of a superhero correlated to the number of years since he joined the team?

Again, I filtered out the Avengers with invalid Year values. Then, I plotted a point for each Avenger based on the two dimensions. Based on the points it is possible to observe a hard limit of 52 on the number of years.since.joining, which is natural considering the first team of Avengers was formed in 1963.

I also plotted a "locally estimated scatterplot smoothing" with geom_smooth(). 

```{r}
  avengers_tidydf %>%  
  filter(Year != 1900) %>%
  ggplot(aes(x= Years.since.joining, y = Appearances)) +
  geom_jitter() + 
  geom_smooth(method = loess) + 
  scale_y_continuous(name = "Appeearances") +
  scale_x_continuous(name = "Number of years since joining") 
```


This plot shows a positive correlation between these two variables, but there might be some abnormality because of more popular superheros, such as Spider-Man. By filtering out the Avengers above 4000, 2000, 1000, and 500 appearances, we can see that Avengers between 10-20 years experience some loss of popularity, while newer and older Avengers see a positive correlation between these two variables.

```{r}
library(gridExtra)
scatterplotAppereancesYearsSinceJoining<- function (n=5000, m = 60){
  avengers_tidydf %>%  
  filter(Year != 1900, Appearances<n ,Years.since.joining < m) %>%
  ggplot(aes(x= Years.since.joining, y = Appearances)) +
  geom_jitter() + 
  geom_smooth(method = loess) +
  ggtitle(paste("Apperances<", n, "; Years<", m)) +
  scale_y_continuous(limits = c(-1,n), name = "Appeearances") +
  scale_x_continuous(limits = c(-1,m), name = "Number of years since joining") 
}

lapply(list(4000,2000,1000,500), function(n){scatterplotAppereancesYearsSinceJoining(n)}) %>%
  grid.arrange(ncol = 2,grobs=.);

```

